JD notes on effects

### 2022-07-05 manuscript/abstract.tex
trying very hard ☺
No major progress, we might need to discuss first.
Handing over to JD after meeting

% Title must be 250 characters or less.
\begin{flushleft}
{\Large
\textbf\newline{Describing the curves: uncertainty estimation and bias correction for predictions in simple and generalized linear (mixed) models} % Please use "sentence case" for title and headings (capitalize only the first word in a title (or heading), the first word in a subtitle (or subheading), and any proper nouns).
}
\newline
% Insert author names, affiliations and corresponding author email (do not include titles, positions, or degrees).
\\
Steve Cygu\textsuperscript{1},
Benjamin M. Bolker\textsuperscript{1,2},
Jonathan Dushoff\textsuperscript{1,2}
\\
\bigskip
\textbf{1} School of Computational Science and Engineering, McMaster University, Hamilton, Ontario, Canada
\\
\textbf{2} Department of Biology, McMaster University, Hamilton, Ontario, Canada
\\
\bigskip

@@

>>>>

% Title must be 250 characters or less.
\begin{flushleft}
{\Large
\textbf\newline{Outcome plots: uncertainty estimation and bias correction for predictions and effects in simple and generalized linear (mixed) models} % Please use "sentence case" for title and headings (capitalize only the first word in a title (or heading), the first word in a subtitle (or subheading), and any proper nouns).
}
\newline
% Insert author names, affiliations and corresponding author email (do not include titles, positions, or degrees).
\\
Steve Cygu\textsuperscript{1},
Benjamin M. Bolker\textsuperscript{1,2},
Jonathan Dushoff\textsuperscript{1,2}
\\
\bigskip
\textbf{1} School of Computational Science and Engineering, McMaster University, Hamilton, Ontario, Canada
\\
\textbf{2} Department of Biology, McMaster University, Hamilton, Ontario, Canada
\\
\bigskip

@@

\jd{This talks about a lot of terms that we don't define. Remember to rewrite!}
\fix{BC to rewrite}

### 2022-07-05 manuscript/body.tex
Handing over to JD after meeting

\item \textbf{Model center:} A point corresponding to a column-wise mean of the model matrix (the mean of one or more model variables). Also referred to as center point. We refer to the prediction approach that uses model center as mean-based. The center point for a set of model variables corresponding to an input variable may not represent a possible value of the input variable -- the case in Fig~\ref{fig:justify_plots}B. In simple linear models (without) interaction, the mean of the input variable can be used as a proxy for the center point. If there are more than one model variables associated with the focal input variable, however, the model center does not correspond to any given value of the variable, and we do not necessarily get the ``nice'' crossing we get in simpler cases, see Fig~\ref{fig:pred_cubic_plots}.
\item \textbf{Reference point:} Corresponds to the value or values chosen for non-focal input variables, when estimating the predictions and effects. Typically the center point (mean-based), but can instead be a whole-sample of quantiles or observations. Appropriately choosing the reference point is very crucial in generating correct predictions (for bias correction) especially when dealing with non-linear link functions.
\item \textbf{Anchor:} The value chosen for the focal input variable when estimating effect confidence intervals. The anchor choice does not affect the main prediction estimates. Typically chosen as the center point of the predictor variables corresponding to the focal input variable but other sensible values can be chosen too. If we are trying to focus on effects, the anchor from which we calculate confidence intervals becomes important. This is not true for the more conventional prediction confidence intervals. We can choose any anchor, but the center point is a natural, stable choice, we call this.

@@

To illustrate mean-based approach, consider a simple linear model with linear predictor $\eta = \bX\bbeta$ and let $g(\boldmu) = \boldeta$ be an identity link function, where $\boldmuh$ is the expected value of response variable $\hat{y}$. Let $\bbetah$ be the estimate of $\bbeta$, together with the estimated covariance matrix $\Sigma = V(\bbetah)$ of $\bbetah$. Then the quantity $\hat{\boldeta}^\star = \bX^\star\bbetah$ is the \fix{central (outcome) estimate} with respect to the focal input variable in question \citep{fox2009effect}; where $\bX^\star$ is the centered model matrix constructed by averaging the columns of non-focal predictors in model matrix $\bX$, and together with appropriately chosen values of focal predictor.

@@

\jd{Why is whole-sample common? Does anyone do it at all except us?}

>>>>

\item \textbf{Model center:} A point corresponding to a column-wise mean of the model matrix (the mean of one or more model variables). Also referred to as center point. We refer to the prediction approach that uses model center as mean-based. The center point for a set of model variables corresponding to an input variable may not represent a possible value of the input variable -- the case in Fig~\ref{fig:justify_plots}B. In simple linear models (without) interaction, the mean of the input variable can be used as a proxy for the center point. If there are more than one model variables associated with the focal input variable, however, the model center does not correspond to any given value of the focal predictor, and we do not necessarily get the ``nice'' crossing we get in simpler cases, see Fig~\ref{fig:pred_cubic_plots}.
\item \textbf{Reference point:} Corresponds to the value or values chosen for non-focal predictors, when estimating the predictions and effects. Typically the center point (mean-based), but can instead be a whole-sample of quantiles or observations. Appropriately choosing the reference point is very crucial in generating correct predictions (for bias correction) especially when dealing with non-linear link functions.
\item \textbf{Anchor:} The value chosen for the focal predictor when estimating effect confidence intervals. The anchor choice does not affect the central estimates. Typically chosen as the center point of the model variables corresponding to the focal predictor but other sensible values can be chosen too. If we are trying to focus on effects, the anchor from which we calculate confidence intervals becomes important. This is not true for the predictions. We can choose any anchor, but the center point is a natural and stable choice.

@@

To illustrate mean-based approach, consider a simple linear model with linear predictor $\eta = \bX\bbeta$ and let $g(\boldmu) = \boldeta$ be an identity link function, where $\boldmuh$ is the expected value of response variable $\hat{y}$. Let $\bbetah$ be the estimate of $\bbeta$, together with the estimated covariance matrix $\Sigma = V(\bbetah)$ of $\bbetah$. Then the quantity $\hat{\boldeta}^\star = \bX^\star\bbetah$ is the central estimate with respect to the focal predictor in question \citep{fox2009effect}; where $\bX^\star$ is the centered model matrix constructed by averaging the columns of non-focal predictors in model matrix $\bX$, and together with appropriately chosen values of focal predictor.

@@
